---
title: "Time Series Code"
author: "Team 6: Trendsetters"
date: "3/12/2020"
output: html_document
---

## Time Series Analysis ##

```{r message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(vars)
library(tseries)
```

```{r}
corona2 <- read.csv("./Datasets/COVID19_open_line_list.csv")
corona3 <- read.csv("./Datasets/time_series_covid_19_confirmed.csv", header = TRUE)
corona4 <- read.csv("./Datasets/time_series_covid_19_deaths.csv", header = TRUE)
corona5 <- read.csv("./Datasets/time_series_covid_19_recovered.csv", header = TRUE)
corona6 <- read.csv("./Datasets/covid_19_data.csv", header = TRUE)
corona7 <- read.csv("./Datasets/recentcorona.csv", header = TRUE)

corona3 <-  gather(corona3, date, confirmed, X1.22.20:X2.20.20, factor_key=TRUE)
corona3$date <- gsub("X","", corona3$date)
corona3$date <- as.Date(corona3$date, "%m.%d.%Y")
corona3$date <- gsub("0020","2020", corona3$date)
corona3$date <- as.Date(corona3$date, "%Y-%m-%d")

corona4 <-  gather(corona4, date, deaths, X1.22.20:X2.20.20, factor_key=TRUE)
corona4$date <- gsub("X","", corona4$date)
corona4$date <- as.Date(corona4$date, "%m.%d.%Y")
corona4$date <- gsub("0020","2020", corona4$date)
corona4$date <- as.Date(corona4$date, "%Y-%m-%d")

corona5 <-  gather(corona5, date, recovered, X1.22.20:X2.20.20, factor_key=TRUE)
corona5$date <- gsub("X","", corona5$date)
corona5$date <- as.Date(corona5$date, "%m.%d.%Y")
corona5$date <- gsub("0020","2020", corona5$date)
corona5$date <- as.Date(corona5$date, "%Y-%m-%d")

corona6$ObservationDate <- as.Date(corona6$ObservationDate, "%m/%d/%Y")

corona3$deaths <- corona4$deaths
corona3$recovered <- corona5$recovered

world <- corona6 %>% group_by(ObservationDate) %>% mutate(sum_deaths = sum(Deaths), sum_confirmed = sum(Confirmed), sum_recovered = sum(Recovered))
world <- world %>% filter(Country.Region == "South Korea")
world %>%
  ggplot(aes(x = ObservationDate)) + 
  geom_line(aes(y = sum_confirmed, colour = "confirmed")) + 
  geom_line(aes(y = sum_deaths, colour = "deaths")) +
  geom_line(aes(y = sum_recovered, colour ="recovered")) +
  scale_color_manual(values = c(confirmed="orange", deaths="red", recovered="green" )) +
  labs(color="Status") +
  xlab('Time') +
  ylab('Count') +
  ggtitle("Global Corona Infection count")
corona2$country <- as.character(corona2$country)
```

```{r}
ct <- world[,c("ObservationDate", "sum_deaths", "sum_confirmed", "sum_recovered")]

plot.ts(ct)

#Divide into train and test
inds <- seq(as.Date("2020-01-22"), as.Date("2020-03-02"), by = "day")
timedata = ts(ct[,-1], start = c(2020, as.numeric(format(inds[1], "%j"))),frequency=365)
plot.ts(timedata)
time.c <- timedata[,2]
time.d <- timedata[,1]
time.r <- timedata[,3]
```

```{r}
###############
#Regular differencing for confirmed

reg.diff=diff(time.c, lag=1,diff=1)

acf(reg.diff,lag=50)

# conclusion : regular differencing is enough for confirmed
c.diff <- reg.diff

###############
#Regular differencing for recovered

reg.diff=diff(time.r, lag=1,diff=1)

acf(reg.diff,lag=50)

#Seasonal differencing for recovered

seas.diff=diff(time.r, lag=7,diff=1)

acf(seas.diff,lag=50)


#Regular seasonal differencing for recovered

seas.reg.diff=diff(reg.diff, lag=7,diff=1)

acf(seas.reg.diff,lag=50)


##See acf plot of each differencing method

par(mfrow=c(3,1))

acf(reg.diff,lag=50, main="reg only diff")

acf(seas.diff,lag=50,main="seas only diff")

acf(seas.reg.diff,lag=50,main="reg and seas diff")

# when convinced, type 

dev.off() 
# conclusion : regular seasonal differencing is enough for recovered
r.diff <- seas.reg.diff

#Regular differencing for deaths

reg.diff=diff(time.d, lag=1,diff=1)

acf(reg.diff,lag=50)

#Seasonal differencing for deaths

seas.diff=diff(time.d, lag=7,diff=1)

acf(seas.diff,lag=50)

#Regular seasonal differencing for death

seas.reg.diff=diff(reg.diff, lag=7,diff=1)

acf(seas.reg.diff,lag=50)


##See acf plot of each differencing method

par(mfrow=c(3,1))

acf(reg.diff,lag=50, main="reg only diff")

acf(seas.diff,lag=50,main="seas only diff")

acf(seas.reg.diff,lag=50,main="reg and seas diff")

# when convinced, type 

dev.off() 

# conclusion : regular seasonal differencing is enough for death
d.diff <- seas.reg.diff
```

### Time Series Modeling ###
```{r}
#ARIMA Model
par(mfrow = c(2,1))
acf(time.c)
pacf(time.c)
dev.off()

acf(c.diff)

model.c=arima(time.c,order=c(1,1,0), seas=list(order=c(0,0,0),7))
par(mfrow = c(3,1))
acf(residuals(model.c),main="ACF of residuals of  model.c")
pacf(residuals(model.c),main="ACF of residuals of  model.c")
acf(residuals(model.c)^2,main="ACF of squared residuals of  model.c")
dev.off()
Box.test(residuals(model.c),lag=12,type="Ljung-Box")
model.c$coef

#AIC, and Sigma squared
model.c

#Plotting Arima for confirmed
LT.forecast = predict(model.c,7)$pred

LT.forecast.se = predict(model.c,7)$se

inds <- seq(as.Date("2020-03-3"), as.Date("2020-03-9"), by = "day")
predict.val = ts(LT.forecast, start = c(2020, as.numeric(format(inds[1], "%j"))),frequency=365)
ci.low = ts(LT.forecast - 1.96*LT.forecast.se, start = c(2020, as.numeric(format(inds[1], "%j"))),frequency=365)
ci.high = ts(LT.forecast + 1.96*LT.forecast.se,start = c(2020, as.numeric(format(inds[1], "%j"))),frequency=365)

ts.plot(cbind(time.c, predict.val,ci.low, ci.high), 
        lty=c(1,1,2,2), col=c("black", "red","blue","blue"),
        main = "COVID-19 confirmed patients using Arima",
        ylab = "Confirmed",
        xlab = "Week")
legend("bottomright", legend=c("Confirmed","forecast","Prediction Interval"), lty=c(1,1,2), col=c("black", "red","blue"))

sqrt(sum((LT.forecast- corona7$Confirmed)^2) / length(LT.forecast))

#Exponential smoothing for deaths
es = HoltWinters(time.c, gamma = FALSE)
es

Box.test(residuals(es),lag=12,type="Ljung-Box")
acf(residuals(es))
pacf(residuals(es))
fitted.es = fitted(es)
LT.forecast.es.c= predict(es, n.ahead = 7, prediction.interval = TRUE)
ts.plot(time.c, LT.forecast.es.c,lty=c(1,1,2,2), col=c(1,2,3,3), xlab = "Week", ylab="Weekly Corona Confirmed", main = "COVID-19 Confirm using Exp Smoothing")
legend("bottomright", legend=c("Confirmed","forecast","Prediction Interval"), lty=c(1,1,2), col=1:3)

sqrt(sum((LT.forecast.es.c[,1]- corona7$Confirmed)^2) / length(LT.forecast.es.c))
```

\[
  \begin{aligned}
  y_t = \gamma_{0t} + \gamma_{1,t}t + w_t  \\
  \hat{\gamma_{0t}} = \hat{\gamma_{0,t-1}} + 1(y_{t-1} - \hat{\gamma_{0,t-1}}) + \hat{\gamma_{1,t-1}} \\
  \hat{\gamma_{1,t}} = 0.262(\hat{\gamma_{0t}} - \hat{\gamma_{0,t-1}}) + (1 - 0.262)\hat{\gamma_{1,t-1}}
  \end{aligned}
  \]

```{r}
#Exponential smoothing for recovered
es = HoltWinters(time.r, gamma = FALSE)
es

Box.test(residuals(es),lag=12,type="Ljung-Box")

plot(es)
fitted.es = fitted(es)
plot(fitted.es)
LT.forecast.es.r= predict(es, n.ahead = 7, prediction.interval = TRUE)
ts.plot(time.r, LT.forecast.es.r,lty=c(1,1,2,2), col=c(1,2,3,3), xlab = "Week", ylab="Weekly Corona confirmed", main = "COVID-19 recovered patients using Exp Smoothing")
legend("bottomright", legend=c("Confirmed","forecast","Prediction Interval"), lty=c(1,1,2), col=1:3)

sqrt(sum((LT.forecast.es.r[,1]- corona7$Recovered)^2) / length(LT.forecast.es.r))
```

\[
  \begin{aligned}
  y_t = \gamma_{0t} + \gamma_{1,t}t + w_t \\
  \hat{\gamma_{0t}} = \hat{\gamma_{0,t-1}} + 0.487(y_{t-1} - \hat{\gamma_{0,t-1}}) + \hat{\gamma_{1,t-1}} \\
  \hat{\gamma_{1,t}} = (\hat{\gamma_{0t}} - \hat{\gamma_{0,t-1}})
  \end{aligned}
  \]

```{r}
#Exponential smoothing for deaths

es = HoltWinters(time.d, gamma = FALSE)
es

Box.test(residuals(es),lag=12,type="Ljung-Box")

plot(es)
fitted.es = fitted(es)
plot(fitted.es)
LT.forecast.es.d= predict(es, n.ahead = 7, prediction.interval = TRUE)
ts.plot(time.d, LT.forecast.es.d,lty=c(1,1,2,2), col=c(1,2,3,3), xlab = "Week", ylab="Weekly Corona death", main = "COVID-19 death using Exp Smoothing")
legend("bottomright", legend=c("Death","forecast","Prediction Interval"), lty=c(1,1,2), col=1:3)

sqrt(sum((LT.forecast.es.d[,1]- corona7$Death)^2) / length(LT.forecast.es.d))
```

$$
  \begin{aligned}
y_t = \gamma_{0t} + \gamma_{1,t}t + w_t  \\
\hat{\gamma_{0t}} = \hat{\gamma_{0,t-1}} + 0.572(y_{t-1} - \hat{\gamma_{0,t-1}}) + \hat{\gamma_{1,t-1}} \\
\hat{\gamma_{1,t}} = 0.536(\hat{\gamma_{0t}} - \hat{\gamma_{0,t-1}}) + (1 - 0.536)\hat{\gamma_{1,t-1}}
\end{aligned}
$$
  
To achieve stationarity on all the data, I used normal and seasonal differencing to erase the autocorrelation between residuals. P-values from all the Ljung Box test were 0.9987, 0.9512, 0.4011 which stayed in the H0 of no autocorrelation. RMSE of Predicted values from Confirmed count was 3936.479, Recover count was 914.5039, and Death count was 150.4928.

```{r}
ts.plot(cbind(time.c, time.d, time.r), cbind(LT.forecast.es.c[,1], LT.forecast.es.d[,1], LT.forecast.es.r[,1]),lty=c(1,1,1,2,2,2), col=c(1,2,3,4,4,4), xlab = "Week", ylab="Patient Count", main = "COVID-19 prediction using Exp Smoothing")
legend("topleft", legend=c("Confirmed","Death", "Recovered", "Prediction"), lty=c(1,1,1,2), col=1:4)
```
